### 2019-02-28 manuscript/methods.tex
first draft of methods and results for pathways

@@

% \subsection{Pathways to polyploidy} % E todo
% robertson2011 method, and then with diversification
% propogation matrix, stable stage structure
% avoid "probability" and "transition matrix"
% analogous to Leslie matrix?

>>>>

\subsection{Pathways to polyploidy}

Considering ploidy and breeding system together, there are two evolutionary pathways from SI diploid to SC polyploid \citep{brunet2001, robertson_2011}.
In the one-step pathway, the CP state is produced directly from the ID state when whole genome duplication disables SI.
In the two-step pathway, the CD state is an intermediate: SI is first lost, and later the SC diploid undergoes polyploidization.
We quantify the relative contribution of these pathways to polyploidy in two ways, each using the MAP rate estimates from the IC/CD/CP model.

Both of our methods are based on a propogation matrix that describes flow from ID to CP, as in \citet{robertson_2011}.
We insert an artificial division in the CP state, so that one substate contains the CP species that arrived via the one-step pathway and the other substate contains the CP species that arrived via the two-step pathway.
We consider only unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and not strong support for diploidization.
First, we consider only the rates of transitions between these states, placing them in the propogation matrix \myvec{Q}.
The matrix $\myvec{P} = \exp(\myvec{Q} t)$ then provides the probabilities of changing from one state to any other state after time $t$.
Closed-form solutions for the two pathway probabilities are provided in \citet{robertson_2011}.
Our results will differ from those of \citet{robertson_2011} because our transition rate estimates come from a dated phylogeny and a model that allows for state-dependent diversification.
%
Second, we consider not only transitions between states but also diversification within each state.
State-dependent diversification can change the relative contributions of the two pathways.
In particular, if the net diversification rate is small for CD, the two-step pathway will contribute relatively less.
We therefore include the difference between speciation and extinction along the diagonal elements of the propogation matrix.
As before matrix exponentiation provides the relative chance of changing from one state to any other state after time $t$.
In this case, however, these are not probabilities because diversification changes the number of lineages as time passes.
We can still use their ratios to consider the relative contribution of each pathway, though, analogous to the normalized age structure in a growing population.

% E todo: for the case with net div, write out the propogation matrix and equations for pathway contributions

@@

### 2019-02-28 manuscript/results.tex
notes while talking to Rosana
- a variety of small comment deletions, comments and changes (mainly formatting, etc.)
first draft of methods and results for pathways

@@

%B: how about below; was: We found that when considered alone, ploidy and breeding system each showed significant associations with net diversification differences. 
Individual trait-dependent diversification inference, of ploidy and breeding system considered separately, each showed significant associations with net diversification differences. 
When considered together, however, the effect of breeding system dominated the effect of ploidy, although hidden factors played an important role as well.

@@

% E: I moved the discussion of \rho_I and \rho_C values into the Pathways section.

@@

% E: If we're going to include pathway results for real, I need to write more in this section and incorporate the figure.  TODO

Polyploidization rates from self-compatible diploid $\rho_{CD}$ and from self-incompatible diploid $\rho_{ID}$ are not significantly different (see supplementary information figures).
However, the distribution of $\rho_{ID}$ seems to always consider faster values compare to the distribution of $\rho_{CD}$.
This result is highlighted in ancestral reconstructions of the $ID/CD/CP$ model where transitions from ID to CP are more probable than transitions from CD to CP. % Emma here we put the figure?
%B: I am not sure I understand the "However, " sentence. TODO.

@@

    \end{center}
    \caption{
        Median rate estimates for all fitted models. %B: "...all fitted/considered/evaluated models."
        Units are per million years.
        Two comma-separated numbers refer to the $A$ and $B$ hidden states, and --- means the parameter was not present in the model.
    %B: perhaps include this in the table so lazies don't have to go seeking subscripts?
    Net diversification rates ($r$) are subscripted with trait state initials (Diploid, Polyploid, Incompatible, Compatible). Transition rates are $\rho$ (polyploidization), subscripted with background breeding system state; $\delta$ (diploidization); and  $q_{IC}$ (loss of self-incompatibility).
    The upper section is for models without diploidization, and the lower section is for models with diploidization.
    The supplemental figures show the corresponding distributions of parameter estimates.
    }
    \label{table:estimates}
\end{supptable}

@@

@@

@@

\begin{suppfigure}
\includegraphics[width=\textwidth]{hisseDPposteriordist.pdf}
\caption{Posterior distribution for each of the parameters in the D/P+A/B, polyploidy model} % XXX
% fixme: offset plotting error in "relative extinction" panel
\label{suppfigure:DPAB}
\end{suppfigure}

>>>>

% NOTE: will need to update figure/panel refs if making no-delta models primary

@@

Individual trait-dependent diversification inference, of ploidy and breeding system considered separately, each showed significant associations with net diversification differences. 
When considered together, however, the effect of breeding system dominated the effect of ploidy, although hidden factors played an important role as well.

@@

Evolution from SI diploids to SC polyploids can proceed through two different pathways.
Determining the relative contribution of these pathways based on our estimated transition rates from the ID/CD/CP model, we find that the one-step pathway is more likely on short timescales and the two-step pathway is more likely on long timescales (\cref{figure:pathways}, left panels).
When not much time has elapsed, the one-step pathway is more likely because only one event need happen.
When more time has elapsed, the two-step pathway is more likely because the rate of loss of SI within diploids, $q_{IC}$, is greater than the rate of polyploidization for SI species, $\rho_I$ (\cref{suppfigure:IDCDCPnodip}).
These conclusions are the same as those of \citet{robertson_2011} under one of their branch length approximations.

When rates of lineage diversification are considered, however, the conclusions change.
Even over long timescales, the two-step pathway contributes less (\cref{figure:pathways}, right panels).
The lower rate of net diversification in the CD state, relative to ID, means that relatively fewer lineages are available to complete the second step of the two-step pathway.

@@

% R will add a bit about param confidence

@@

    \end{center}
    \caption{
        Median rate estimates for all fitted models.
        Units are per million years.
        Two comma-separated numbers refer to the $A$ and $B$ hidden states, and --- means the parameter was not present in the model.
    %B: perhaps include this in the table so lazies don't have to go seeking subscripts?
    Net diversification rates ($r$) are subscripted with trait state initials (Diploid, Polyploid, Incompatible, Compatible). Transition rates are $\rho$ (polyploidization), subscripted with background breeding system state; $\delta$ (diploidization); and  $q_{IC}$ (loss of self-incompatibility).
    The upper section is for models without diploidization, and the lower section is for models with diploidization.
    The supplemental figures show the corresponding distributions of parameter estimates.
    }
    \label{table:estimates}
\end{supptable}

@@

% TODO: swap panels between the figures so that the first is -delta and the second is +delta

@@

% E todo: could make this look a little less klunky
\begin{figure}
    \centering \includegraphics[width=0.9\textwidth]{pathways}
    \caption{
        Contributions of the two pathways to polyploidy.
        When considering only rates of transitions among the states (left), the one-step pathway dominates on short timescales and the two-step on long timescales.
        When also considering diversification within each state (right), the one-step pathway dominates over any timescale.
        The lower panels plot the ratio of the two pathway contributions in the upper panels.
    }
    \label{figure:pathways}
\end{figure}

@@

\begin{suppfigure}
\includegraphics[width=\textwidth]{hisseDPposteriordist.pdf}
\caption{Posterior distribution for each of the parameters in the D/P+A/B, polyploidy model.  The axis is offset in one location so that the two overlapping distributions can be seen.} % XXX
\label{suppfigure:DPAB}
\end{suppfigure}
